---
title: "Assignment 9"
author: "Suzanne Combs"
output: 
  html_document:
    code_folding: hide
date: "December 2, 2024"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Load in packages ####
library(tidyverse)
library(GGally)
library(modelr)
library(easystats)
library(patchwork)
```
<br></br>
*The codes are included, but hidden. To view the code click the "show" button
<br></br>
<br></br>
For assignment 9, my task is to:

* Use the data set “/Data/GradSchool_Admissions.csv”

* Explore and model the predictors of graduate school admission

* Document data explorations, figures, and conclusions in a reproducible R-markdown report

<br></br>
I'll first load the data set. 
```{r, message=FALSE}
## Use the data set "/Data/GradSchool_Admissions.csv" ####
# Load in the dataset
df <- read_csv("../../Data/GradSchool_Admissions.csv")
```
<br></br>
Then clean it up. To clean the data set, I will change "admit" to a logical because it can either be true or false. I will also change "rank" to a factor since there are only 4 different level options.
<br></br>
```{r}
# Clean the dataset. Changing the variable types
df$admit <- as.logical(df$admit)
df$rank <- as.factor(df$rank)
```
<br></br>
Next, I will explore the data
<br></br>
```{r}
glimpse(df)
```
<br></br>
and view different plots using ggpairs for a quick look
<br></br>
```{r, message=FALSE}
ggpairs(df)
```
<br></br>
<br></br>
There seems to be differences between the chance of being admitted between schools of different ranks. This is one of the things I will look at more in my visualization.
<br></br>
<br></br>
```{r}
# Boxplots
p1 <- ggplot(df, aes(y=admit, x=gre, fill=rank)) +
  geom_boxplot() +
    labs(title = "Admittance based on GRE")

p2 <- ggplot(df, aes(y=admit, x=gpa, fill=rank)) +
  geom_boxplot() +
  labs(title = "Admittance based on GPA")

# Jitter Scatterplots
p3 <- ggplot(df, aes(y=admit, x=gre, color=rank)) +
  geom_jitter() +
  theme_minimal() +
  labs(title = "Admittance based on gre")

p4 <- ggplot(df, aes(y=admit, x=gpa, color=rank)) +
  geom_jitter() +
  theme_minimal() +
  labs(title = "Admittance based on gpa")
p1 + p2
p3 + p4
```
<br></br>
<br></br>
It is difficult to tell if rank, gre, and gpa effects admittance rates based on these plots. I am going to create three different models to see what is significant.

<br></br>
Model 1
```{r}
# Models
mod1 <- glm(data=df, formula = admit ~ gpa:gre + rank, family = "binomial")
summary(mod1)
```
From this model, the interaction between gpa and gre appear to be significant. Rank also seems to be significant, although rank2 is less significant than the other ranks.
<br></br>
Model 2
```{r}
mod2 <- glm(data=df, formula = admit ~ gpa + gre + rank, family = "binomial")
summary(mod2)
```
This model, rank3 and rank4 are significant and gpa, gre, and rank2 are less significant. 
<br></br>
Model 3
```{r}
mod3 <- glm(data=df, formula = admit ~ gpa * gre * rank, family = "binomial")
summary(mod3)
```
This model only shows to be slightly significant for some of the variables.
<br></br>
Now that I have created 3 models, I can compare the models to see which would work best.
<br></br>
Comparing performance
```{r}
# Models
# Compare the performance
comps <- compare_performance(mod1, mod2, mod3, rank=TRUE)
comps
```

Plot the comparisons
```{r}
# Plot the comparisons
comps %>% 
  plot()
```

And look at the mean-squared-error for each model

Model 1:
```{r}
# Mean Residuals (Look for exact name from other assignment)
mean(mod1$residuals^2)
```

Model 2:
```{r}
mean(mod2$residuals^2)
```

Model 3:
```{r}
mean(mod3$residuals^2)
```

With all these comparisons to consider, I am going to choose model 1 because it has a much higher performance score than model 2 or model 3. 
Model 1 had a performance score of 66.67% while Model 2 was 49.69% and Model 3 was 42.19%.

The mean residuals were very similar for model 1 and model 3, but model 1 is a less complicated model than model 3.
<br></br>
<br></br>
I will add predictions from mod1 to a new dataframe
```{r}
# Add predictions from mod1 to a new dataframe
df2 <- add_predictions(df, mod1, type="response")
```

and create plots comparing gre and gpa on admittance
```{r, message=FALSE}
# Predict with model 1 how gre effects admittance
p1 <- ggplot(df2, aes(x=gre, y=pred, color=rank)) + # As gre goes up, the probability goes up
  geom_point() +
  geom_smooth(se=FALSE, method="lm") +
  labs(title="Admittance based on GRE",
       x="GRE",
       y="Prediction")

# Predict with model 1 how gpa effects admittance
p2 <- ggplot(df2, aes(x=gpa, y=pred, color=rank)) + # As gpa goes up, the probability goes up
  geom_point() +
  geom_smooth(se=FALSE, method="lm") +
  labs(title="Admittance based on GPA",
       x="GPA",
       y="Prediction")

p1 + p2
```
<br></br>
From these predictions, it looks like the higher your GRE and GPA the more likely it is to be admitted. Also having a rank 1 school gives you the best chance to be admitted.